---
title: "STAT 231: Problem Set 1B"
author: "Vaibhav Shah"
date: "due by 5 PM on Friday, February 26"
output: pdf_document
---

Series B homework assignments are designed to help you further ingest and practice the material covered in class over the past week(s).  You are encouraged to work with other students, but all code must be written by you and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps1B.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close the course-content repo project in RStudio
\item Open YOUR repo project in RStudio
\item In the ps1B.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope.  Don't forget to select which of your pages are associated with each problem.  \textit{You will not get credit for work on unassigned pages (e.g., if you only selected the first page but your solution spans two pages, you would lose points for any part on the second page that the grader can't see).} 
\end{enumerate}

```{r, setup, include=FALSE}
# Load packages here 
library(mdsr)   
library(tidyverse)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",   # slightly smaller font for code
  message=FALSE)   
```

\newpage 

# If you discussed this assignment with any of your peers, please list who here:

> ANSWER:



\newpage
# MDSR Exercise 2.5 (modified)

Consider the data graphic for Career Paths at Williams College at: https://web.williams.edu/Mathematics/devadoss/careerpath.html.  Focus on the graphic under the "Major-Career" tab.

a. What story does the data graphic tell?  What is the main message that you take away from it?

> ANSWER: You see the distribution of job outcomes for each major and the distribution of majors that went into each job outcome. The graph shows that there tend to be job sectors that majors tend to conglomerate to, but there are people of all majors in each sector and vice-versa. 

b. Can the data graphic be described in terms of the taxonomy presented in this chapter?  If so, list the visual cues, coordinate system, and scale(s).  If not, describe the feature of this data graphic that lies outside of that taxonomy.

> ANSWER: The visual cues of the data graphic can be described with the textbook's taxonomy to some extent, but the graphical, interconnected network and lack of numerical data shown well lie outside the textbook's normal taxonomy for coordinate systems and scales. Every major and job department is given its own color and patial circumference along a circle split into two halves for jobs/majors, and all jobs are connected to the majors that contribute to its workforce while all majors are connecte dot the job departments their graduates venture into. The width of the curved line connected is proportional to the portion of the total major/workforce combination that line represents. We can assume this means the scale is numeric and based on percentages to some extent, however it is not explicitly stated anywhere, and this can only be assumed since the sum of line widths seems to correspond to the width of the partial circumference a major/job occupies. 

c. Critique and/or praise the visualization choices made by the designer.  Do they work? Are they misleading?  Thought-provoking?  Brilliant?  Are there things that you would have done differently?  Justify your response.

> ANSWER: The visualiztion is thought-provoking since it shows how there is correlation between major and job, but the strength of the correlation caries greatly dpeending on what job or major you're looking at. Furthermore, basically everything is connected which shows that your major does not necessarily determine your job path, and everyone in a job department isn't necessarily from a particular major or group of majors. On the other hand, percentages should've been explicitly shown somehow. There was a clear lack of numerical scale of any kind, and the width of lines was hard to follow as they curved, looped, bended, etc. The differentiation of color for nearby jobs/majors was very helpful though in following along. 


\newpage
# Spot the Error (non-textbook problem)

Explain why the following command does not color the data points blue, then write down the command that will turn the points blue.

> ANSWER:The geom_point funciton can affect how and where points are drawn, and the color option in the aes function only differentiates groups by color but isn't used to designate a particular color to all points.

```{r}
library(ggplot2)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))  + scale_color_manual(values = "blue")
```


\newpage
# MDSR Exercise 3.6 (modified)

Use the `MLB_teams` data in the `mdsr` package to create an informative data graphic that illustrates the relationship between winning percentage and payroll in context. What story does your graph tell?

> ANSWER: There is a small correlation between winning percentage and payroll, but it's very weak. Even wehn accounting for year, league, and even the individual team, there isn't a clear, consistent pattern in the relationship between payroll and winning percentage.

```{r}
library(mdsr)
MLB_teams
rel <- ggplot(data = MLB_teams) + geom_point(mapping = aes(x = WPct, y = payroll))

rel + facet_wrap(~lgID)
rel + facet_wrap(~yearID, nrow=3)
rel + facet_wrap(~teamID, nrow=7)
```


\newpage
# MDSR Exercise 3.10 (modified)

Using data from the `nasaweather` package, use the `geom_path()` function to plot the path of each tropical storm in the `storms` data table (use variables `lat` (y-axis!) and `long` (x-axis!)).  Use color to distinguish the storms from one another, and use facetting to plot each `year` in its own panel.  Remove the legend of storm names/colors by adding `scale_color_discrete(guide="none")`. 

Note: be sure you load the `nasaweather` package and use the `storms` dataset from that package!

```{r}
library(nasaweather)
ggplot(data = storms) + geom_path(mapping = aes(y=lat, x = long, color = name)) + scale_color_discrete(guide="none") + facet_wrap(~year, nrow=2)
```


\newpage
# Calendar assignment check-in

For the calendar assignment:

- Identify what questions you are planning to focus on
- Describe two visualizations (type of plot, coordinates, visual cues, etc.) you imagine creating that help address your questions of interest 
- Describe one table (what will the rows be?  what will the columns be?) you imagine creating that helps address your questions of interest

Note that you are not wed to the ideas you record here.  The visualizations and table can change before your final submission.  But, I want to make sure your plan aligns with your questions and that you're on the right track.

> ANSWER:
Does going to the gym make me more or less likely to follow the schedule I preemptively made for afterwards? Are there certain days of the week where I am more/less likely to do as I say I will, and are there certain days where I unexpectedly do things I wasn't planning to more often?
I could use pie charts to compare what I intend to do each day in terms of activites vs. what I actually do. I could also make a split bar graph with percentages for how closely I follow my schedule on the y axis and on the x axis have time after gym reservation and time before gym reservation. A table could also show my distribution of time allocation to each activity across the week by having rows for each day and columns for each acitivity. Cells could store time spent, and two tables could be made for what I plan and what I actually do.


 